### 2023-08-15 docs/presentation/preprocess.Rmd

---
title: "Preprocessing"
output: 
  html_document:
    theme: flatly
    df_print: paged
date: "2023-07-12"
---

Making sense of the survey data.

```{r import, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.retina = 4,
  fig.showtext = TRUE,
  message = FALSE,
  warning = FALSE
)

# all your themes are belong to us
# library(ggthemes)
library(tidyverse)
library(magrittr)
library(showtext)
font_add_google("Fira Sans", "Fira Sans")
```

```{r data}
analysis <- readRDS("../../data/derived/public/analysis_hegs_rpl.rds")
source("data/constants.R")
```

## Groups

Dropping NAs for brevity

```{r drop na}
analysis %<>%
  drop_na(Q3_recoded, Q4_quantqual) %>%
  filter(Q3_recoded != "Other")
```

Question 3: primary subfield of research

```{r q3 groups}
analysis %>%
  group_by(Q3_recoded) %>%
  summarise(n = n())
```

Question 4: type of research

```{r q4 groups}
analysis %>%
  group_by(Q4_quantqual) %>%
  summarise(n = n())
```

## Word cloud

Q6
Q10

```{r DRY wordcloud}
column_to_cloud <- function(
    col, nrows = 30,
    data = analysis,
    my_palette = c(
      "#cdb4db",
      "#6d597a",
      "#b56576",
      "#e56b6f",
      "#eaac8b"
    )) {
  require(tidytext)
  word_counts <- data %>%
    drop_na(all_of(col)) %>%
    unnest_tokens("word", col) %>%
    # https://bookdown.org/psonkin18/berkshire/tokenize.html#text-normalization---stemming-and-lemmatization
    # doesn't work - why is `reproduct` different from `reproduc`?
    # mutate(word = SnowballC::wordStem(word, language = "english")) %>%
    anti_join(stop_words, by = "word") %>%
    count(word) %>%
    arrange(desc(n)) %>%
    slice(1:nrows)

  require(wordcloud2)
  wordcloud2(
    word_counts,
    color = rep_len(
      my_palette,
      nrow(word_counts)
    )
  ) %>%
    return()
}
```

```{r word clouds}
clouds <- map(cloud_cols, column_to_cloud) %>%
  set_names(cloud_cols)
# saveRDS(clouds, file = "data/clouds.rds")
```

```{r q6 cloud}
clouds_q6_q3 <- map(
  subfields, ~ analysis %>%
    filter(Q3_recoded == .x) %>%
    column_to_cloud("Q6_definition", data = .)
) %>%
  set_names(subfields)

clouds_q6_q4 <- map(
  types, ~ analysis %>%
    filter(Q4_quantqual == .x) %>%
    column_to_cloud("Q6_definition", data = .)
) %>%
  set_names(types)

clouds_q6 <- append(clouds_q6_q3, clouds_q6_q4)
clouds_q6$Everyone <- clouds$Q6_definition
saveRDS(clouds_q6, file = "data/clouds_q6.rds")
```

```{r q19 cloud}
clouds_q19_q3 <- map(
  subfields, ~ analysis %>%
    filter(Q3_recoded == .x) %>%
    column_to_cloud("Q19_reason_for_rep", data = .)
) %>%
  set_names(subfields)

clouds_q19_q4 <- map(
  types, ~ analysis %>%
    filter(Q4_quantqual == .x) %>%
    column_to_cloud("Q19_reason_for_rep", data = .)
) %>%
  set_names(types)

clouds_q19 <- append(clouds_q19_q3, clouds_q19_q4)
clouds_q19$Everyone <- clouds$Q19_reason_for_rep
saveRDS(clouds_q19, file = "data/clouds_q19.rds")
```

## Bar charts
Order Q5
```{r}
analysis %<>%
  mutate(across(all_of(four_level_cols), ~ factor(., levels = four_levels)))
```

TODO: Rename Q6 labels
```{r eval=FALSE, include=FALSE}
analysis %<>%
  mutate(label = case_match(
    label,
    "Omit" ~ NA,
    "Standard" ~ "Assess Prior Work",
    "Experiment" ~ "Repeatable Experiments",
    "Data Collection" ~ "Data Collection",
    "Transparency/Extension" ~ "Transparency/Extension"
  ))
```

## Q7 Likert
```{r}
q7_to_likert <- function(data = analysis) {
  data %>%
    # make data for likert
    pivot_longer(cols = starts_with("Q7_value_")) %>%
    group_by(name, value) %>%
    summarise(n = n()) %>%
    mutate(value = factor(value, levels = agree_levels)) %>%
    replace_na(list(value = "Don't know")) %>%
    group_by(name) %>%
    mutate(perc = n / sum(n)) %>%
    mutate(
      yeah = ifelse(any(value == "Agree"), perc[value == "Agree"], 0),
      idk = ifelse(any(value == "Don't know"), perc[value == "Don't know"] / 2, 0),
      nah = ifelse(any(value == "Disagree"), perc[value == "Disagree"], 0),
      start = case_match(
        value,
        "Strongly agree" ~ yeah + idk,
        "Agree" ~ idk,
        "Don't know" ~ -idk,
        "Disagree" ~ -idk - perc,
        "Strongly disagree" ~ -idk - nah - perc,
      )
    ) %>%
    ungroup() %>%
    dplyr::select(-c(yeah, nah, idk))
}
```

```{r Q8}
q7_q3 <- map(
  subfields, ~ analysis %>%
    filter(Q3_recoded == .x) %>%
    q7_to_likert()
) %>%
  set_names(subfields)

q7_q4 <- map(
  types, ~ analysis %>%
    filter(Q4_quantqual == .x) %>%
    q7_to_likert()
) %>%
  set_names(types)

q7 <- append(q7_q3, q7_q4)
q7$Everyone <- q7_to_likert()
saveRDS(q7, file = "data/q7.rds")
```

## Save
```{r}
analysis %>%
  saveRDS("data/analysis.rds")
```